Spin canonical divisors of log stacky curves

### 2015-07-01 stacky_write_up/spin_cring.tex
mid intro edits

@@

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} group, i.e. a discrete subgroup of
$PSL_2(\BR)$ acting on the upper half plane $\BH$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_k M_k(\Gamma)$. Often, it is useful to
bound the degree of the generators and relations for $M(\Gamma)$
for general Fuchsian groups $\Gamma$. In the special case that $\Gamma$
has no odd weight modular forms, Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky} give nice bounds.
This work extends their work to all Fuchsian groups $\Gamma$.

We can now consider the orbifold $\Gamma \backslash \BH$ over $\BC$.
For example, in the case $\Gamma$ acts freely on $\BH$, $\Gamma
\backslash \BH$ is a Riemann Surface over $\BC$. It may well be
that $\Gamma \backslash \BH$ is not compact, but we can form a
space $\Gamma \backslash \BH^*$ which is a compactification of
$\Gamma \backslash \BH,$ with associated divisor of cusps $\Delta$.
In order to find generators and relations for $M(\Gamma)$, we
translate the seemingly analytic question of understanding the ring
of modular forms into the algebraic category, using a
generalization of the GAGA principle. By \cite[Proposition 6.1.5]
{vzb:stacky}, there is an equivalence of categories between orbifold
curves and log stacky curves over $\BC$. Hence, for the remainder
of the paper, we will work in the algebraic category.

Let $X$ be a smooth algebraic curve over $k$ of genus $g$. It is
well known that the canonical sheaf $\Omega _X,$ with associated
canonical divisor $K_X$  determines the {\bf canonical map } $\pi:
X \rightarrow \BP_k^{g - 1}$. Then, the {\bf canonical ring} is
defined to be
\begin{align*}
	R(X,K_X) = \bigoplus_{d \geq 0} H^0(X, dK_X),
\end{align*}

\noindent
with multiplication structure corresponding to tensor product of
sections. In the case that $g \geq 2$, $\Omega_X$ is ample, so $X
\cong \proj R$. When $g \geq 2$, Petri's theorem gives generators
and relations for $R(X, K_X)$ showing that in most cases $R(X, K_X)$
is generated in degree 1 with relations in degree 2
(demonstrations are given by Saint-Donat \cite[p. 157]
{saint-donat:proj} and Arbarello--Cornalba--Griffiths--Harris
\cite[Section 3.3]{acgh:algebraic-curves}). This has the pleasant 
geometric consequence that all such canonically embedded curves are 
scheme theoretically cut out by quadratic equations.

Following Voight and Zureick-Brown \cite{vzb:stacky}, we generalize 
Petri's theorem in the direction of stacky curves equipped with
log divisors. For a stacky curve $\sx$ with coarse space $X$ and
stacky points (also called ``fractional points'') $P_1, \ldots, P_r$
with stabilizer orders $e_1, \ldots, e_r \in \BZ_{\geq 2}$,

@@

Then a {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$
where $\Delta \in \di X$ is a log divisor and $2\halfcan \sim K_X +
\Delta + \sum_{i = 1}^{r} \subhalf{e_i} P_i$. The central object of
study in this paper is the {\bf log spin canonical ring} of $(\sx,
\Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(X, d \halfcan).
\end{align*}

\noindent
A brief overview of stacky curves, log divisors, and log spin
canonical rings and references are given in Subsection
~\ref{ssec:stacky-background}.

@@

\begin{rem}
\label{rem:stack-formalism}
Stacky curves may be formalized in the language of stacks (e.g. as
done in the works of Voight and Zureick-Brown \cite{vzb:stacky},
Abramovich and Vistoli \cite{abramovich-vistoli:compactifying}, and
Behrend and Noohi \cite{behrend-noohi:uniformization}), but for the
purposes of this paper, it suffices to treat $\sx$ as a scheme
whose associated divisor group allows a finite number of stacky
points. The remainder of the paper will primarily use the language
of Weil divisors.
\end{rem}

@@

Throughout this paper, stacky curves will be assumed to be tame and
$k$ will always be algebraically closed. We also assume that the
stacky curves are separably rooted, i.e. the residue field of each
stacky points is separable. As noted in Remark
~\ref{rem:stack-formalism}, the language of stacks will be omitted
in this paper, as it is sufficient to treat stacky curves and
related objects $\sx$ as a scheme whose associated divisor group
allows for a finite number of \textbf{stacky points} which may be
viewed heuristically as ``fractional points''. See Voight and
Zureick-Brown for more thorough treatment of these definitions
\cite[Chapter 5]{vzb:stacky}.

@@

\section{High Genus}
\label{sec:g_high}
We now consider the case when the genus is at least 2. In this case, we are able to bound the degrees of generators of $R_\halfcan$ and its ideal of relations. In this section, we do not obtain explicit presentations of $R_\halfcan.$  This contrasts Sections ~\ref{sec:g_1} and ~\ref{sec:g_0} where we not only obtain bounds, but also obtain inductive presentations. The tradeoff is that in the genus 0 and 1 cases, we have to account for with explicit base cases, whereas in this section we apply general results.

@@

The main result of this subsection is that for a log spin curve with no stacky points $(X,\Delta,L)$, the spin canonical ring $R_L$ is generated in degree at most $5$, with relations in degree at most $10$. The case that $\Delta = 0$ was completed by Reid \cite[Theorem 3.4]{reid:infinitesimal}. For $\Delta > 0$, the generation bound is shown in Lemma ~\ref{lem:generation-5} and the relations bound is shown in Lemma ~\ref{prop:relation-10}. Thoughout this subsection, we will implicitly use Remark ~\ref{rem:delta-not-1}, which implies $\deg 2L \geq 2g$ so $H^0(X, 2L)$ is basepoint free by Riemann-Roch. We summarize the results of this subsection in Corollary ~\ref{cor:g-2-presentation-bound}.

The proofs of this subsection results are similar to those in Neves ~\cite[Proposition III.4 and Proposition III.12]{neves:halfcan}. However, the statements differ, as we assume $\Delta > 0$ instead of $\Delta = 0$ and do not assume there is a basepoint free pencil contained in $H^0(X,L).$

@@

We wish to show $f$ is surjective.
Note that $\dim_k \ker f = \dim_k H^0((n-2)L) = (n-3)(g - 1+\frac{\delta}{2})$ using Riemann--Roch and the assumption $n \geq 4.$
Additionally, $\dim_k V \otimes H^0(nL) = 2 \cdot (n- 1)(g - 1+\frac{\delta}{2})$, again using Riemann--Roch.
Therefore,
\begin{align*}
	\dim_k \im f 	& = 2 \cdot (n- 1)(g - 1+\frac{\delta}{2}) -(n-3)(g - 1+\frac{\delta}{2})\\
				& = (n+ 1)(g - 1+\frac{\delta}{2}) = \dim_k H^0((n+2)L).
\end{align*}
Ergo, $f$ is surjective.
\end{proof}

@@

The next step is to bound the degrees of the relations of $R_L$ when $\Delta > 0$. This is done in Proposition ~\ref{prop:relation-10}
%The idea of the proof of Proposition ~\ref{prop:relation-10} is to 
by using the basepoint-free pencil trick to show that if a relation lies in a sufficiently high degree, it lies in the ideal generated by the relations in lower degrees.  In Definition ~\ref{defn:lower-ideal}, we fix notation for the ideal generated by lower degrees relations:

>>>>

\todo{Make the sentences in the abstract shorter}
\todo{Make sure we call all divisors spin canonical divisors}

@@

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} group, i.e. a discrete subgroup of
$PSL_2(\BR)$ acting on the upper half plane $\BH$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_{k=0}^\infty M_k(\Gamma)$. One of the best ways to understand the ring of modular forms is to give a presentation of $M(\Gamma)$
for general Fuchsian groups $\Gamma.$ In order to do so, it is extermely useful to have a bound on the degrees in which the generators and relations can occur. In the special case that $\Gamma$
has no odd weight modular forms, Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky} give tight bounds.
This work extends their work to all Fuchsian groups $\Gamma$.

We can now consider the orbifold $\Gamma \backslash \BH$ over $\BC$.
For example, in the case $\Gamma$ acts freely on $\BH$, $\Gamma
\backslash \BH$ is a Riemann Surface over $\BC$. It may be
that $\Gamma \backslash \BH$ is not compact, but we can form a
compact Riemann Surface, $\Gamma \backslash \BH^*,$ by adding in cusps, with associated divisor of cusps $\Delta$.
In order to find generators and relations for $M(\Gamma)$, we
translate the seemingly analytic question of understanding the ring
of modular forms into the algebraic category, using a
generalization of the GAGA principle. By \cite[Proposition 6.1.5]
{vzb:stacky}, there is an equivalence of categories between orbifold
curves and log stacky curves over $\BC$. Hence, for the remainder
of the paper, we will work in the algebraic category.

Let $X$ be a smooth algebraic curve of genus $g$ over a field $k$. It is
well known that the canonical sheaf $\Omega _X,$ with associated
canonical divisor $K_X$, determines the {\bf canonical map } $\pi:
X \rightarrow \BP_k^{g - 1}$. Then, the {\bf canonical ring} is
defined to be
\begin{align*}
	R(X,K_X) = \bigoplus_{d \geq 0} H^0(X, dK_X),
\end{align*}

\noindent
with multiplication structure corresponding to tensor product of
sections. In the case that $g \geq 2$, $\Omega_X$ is ample, so $X
\cong \proj R$. When $g \geq 2$, Petri's theorem gives generators
and relations for $R(X, K_X)$ showing that in most cases $R(X, K_X)$
is generated in degree 1 with relations in degree 2, as shown in
\cite[p. 157]{saint-donat:proj} and Arbarello--Cornalba--Griffiths--Harris
\cite[Section 3.3]{acgh:algebraic-curves}. This has the pleasant 
geometric consequence that all such canonically embedded curves are
scheme theoretically cut out by degree 2 equations.

Following Voight and Zureick-Brown \cite{vzb:stacky}, we generalize 
Petri's theorem in the direction of stacky curves equipped with
log divisors. For a stacky curve $\sx$ with coarse space $X$ and
stacky points (also called ``fractional points'') $P_1, \ldots, P_r$
with stabilizer orders $e_1, \ldots, e_r \in \BZ_{\geq 2}$, we define

@@

Then a {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$
where $\Delta \in \di X$ is a log divisor and $2\halfcan \sim K_X +
\Delta + \sum_{i = 1}^{r} \subhalf{e_i} P_i$. The central object of
study in this paper is the {\bf log spin canonical ring} of $(\sx,
\Delta, \halfcan),$ notated as
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(X, d \halfcan).
\end{align*}

\noindent
A brief overview of stacky curves, log divisors, and log spin
canonical rings is given in Subsection
~\ref{ssec:stacky-background}.

@@

@@

\begin{rem}
\label{rem:stack-formalism}
Stacky curves may be formally defined in the language of stacks, (as
is done in the works of Voight and Zureick-Brown \cite{vzb:stacky},
Abramovich and Vistoli \cite{abramovich-vistoli:compactifying}, and
Behrend and Noohi \cite{behrend-noohi:uniformization},) but for the
purposes of this paper, it suffices to treat $\sx$ as a scheme
whose associated divisor group has a finite number of stacky
points.

When viewed in the formalism of stacks, all stacky curves will be assumed to be tame and
$k$ will always be algebraically closed. We also assume that the
stacky curves are separably rooted, i.e. the residue field of each
of the stacky points is separable.

With the close relation to stacks in mind, for simplicity, the remainder of this paper is primarily phrased using the language
of Weil divisors.
\end{rem}

@@

\section{High Genus}
\label{sec:g_high}
We now consider the case when the genus is at least 2. In this case, we are able to bound the degrees of generators of $R_\halfcan$ and its ideal of relations. In this section, we do not obtain explicit presentations of $R_\halfcan.$  In contrast, in Section ~\ref{sec:g_1} and Section ~\ref{sec:g_0} we will not only obtain bounds on the generators and relations, but also obtain presentations inductively. The tradeoff is that in the cases of genus 0 and 1, we will have to explicitly deal with base cases, making the induction more involved, while in this section we will be able to appeal to general theory.

@@

The main result of this subsection is that for a log spin curve with no stacky points $(X,\Delta,L)$, the spin canonical ring $R_L$ is generated in degree at most $5$, with relations in degree at most $10$. The case that $\Delta = 0$ was completed by Reid \cite[Theorem 3.4]{reid:infinitesimal}. For $\Delta > 0$, the generation bound is shown in Lemma ~\ref{lem:generation-5} and the relations bound is shown in Lemma ~\ref{prop:relation-10}. Thoughout this subsection, we will implicitly use Remark ~\ref{rem:delta-not-1}, which implies $\deg 2L \geq 2g$ and so $H^0(X, 2L)$ is basepoint free by Riemann-Roch. We summarize the results of this subsection in Corollary ~\ref{cor:g-2-presentation-bound}.

The proofs of this subsection results are similar to those in Neves ~\cite[Proposition III.4 and Proposition III.12]{neves:halfcan}. However, the statements are different, as we assume $\Delta > 0$ instead of $\Delta = 0$ and do not assume there is a basepoint free pencil contained in $H^0(X,L).$

@@

We wish to show $f$ is surjective.
Note that $\dim_k \ker f = \dim_k H^0((n-2)L) = (n-3)(g - 1)$ using Riemann--Roch and the assumption $n \geq 4.$
Additionally, $\dim_k V \otimes H^0(nL) = 2 \cdot (n- 1)(g - 1)$, again using Riemann--Roch.
By rank-nullity, $$\dim_k \im f = 2 \cdot (n- 1)(g - 1) -(n-3)(g - 1) = (n+ 1)(g - 1) = \dim_k H^0((n+2)L).$$
Ergo, $f$ is surjective.
\end{proof}

@@

The next step is to bound the degrees of the relations of $R_L$ when $\Delta > 0$. This is done in Proposition ~\ref{prop:relation-10}. The idea of the proof of Proposition ~\ref{prop:relation-10} is to use the basepoint-free pencil trick to show that if a relation lies in a sufficiently high degree, it lies in the ideal generated by the relations in lower degrees. We now fix notation for this ideal generated by lower degrees relations.
